Fast-Match: Fast and robust matching on large images

### 2014-06-17 paper.tex
New timings figure

%===========================================================
\title{Have your cake and eat it too: Fast robust matching with SIFT features} % Replace with your title
\titlerunning{ACCV-14 submission ID \ACCV14SubNumber}
\authorrunning{ACCV-14 submission ID \ACCV14SubNumber}

@@

\begin{abstract}
\end{abstract}

\section{Introduction}
%
Whenever we match local image features we are faced with a choice between performance and precision. On one hand SIFT features proposed by Lowe \cite{lowe2004sift} have shown again and again to compare favorably to other local image descriptors especially under unrestrained conditions such as perspective change with non-planar scenes \cite{mikolajczyk2005performance,moreels2007evaluation,heinly2012comparative}. On the other, SIFT keypoints and descriptors are slow to compute, the main raison d'\^{e}tre for later local image features. In many applications of computer vision we would like the increase the computational performance in order to work on larger images, bigger image sets, at faster frame rates or with more limited hardware but we cannot give up the additional precision that SIFT affords us over other local features without negative effects for our application.

In this paper we introduce a matching algorithm designed to match features between two images only in image areas that are likely to correspond. This approach is much faster than traditional methods because there is no need to compute descriptors for areas in the image that are not matched. We provide two variations of the algorithm: a \emph{general} and a \emph{retrieval} approach. The \emph{general} approach functions as a traditional matching algorithm and matches two unknown images albeit faster and more robustly than conventional methods. The \emph{retrieval} approach on the other hand assumes that we know one of the images we intend to match beforehand but given this trade-off this approach offers performance a magnitude faster than existing matching methods without compromising on accuracy. We will unimaginatively refer to the proposed algorithm as \emph{Fast-Match} in this paper and make clear from the context whether this refers to the \emph{retrieval} or \emph{general} variation. As an aside we adopt an artificial distinction between the words \emph{match} and \emph{correspondence} to make it clear when a match between two images is correct. The word \emph{match} will be used to denote a result from a matching algorithm that may or may not be correct while a \emph{correspondence} refers to an actual visual correspondence between two points in two images.

@@

Efforts to reduce the computational costs of finding nearest neighbours to feature points has largely been focused on metric trees. Na\"ively the set of nearest neighbours between features in two images can computed by brute force in $O(n^2)$ where $n$ is the number of features in the two images. When SIFT was originally published Lowe proposed using the Best-Bin-First method to approximately search for nearest neighbours \cite{beis1997shape,lowe1999object}. This reduces the computational complexity to $O(nlog(n))$ but even approximate metric trees are hard pressed to compete with brute force due to the high dimensionality of SIFT and the constant costs incurred with constructing and searching in a metric tree. Later work by Muja and Lowe has focused on improving approximate nearest neighbour searches by using several KD-Trees simultaneously while optimizing the tree structure using K-Means to cluster similar features \cite{muja2009fast}. Recent work on knn-graphs shows a lot of promise for high dimensional cases \cite{dong2011efficient}. We later review these improvements and their effect on efficiently matching large scale images.

@@

Brown and Lowe \cite{brown2005multi} extend ratio match to deal with a set of images by using not the ratio of the best and second best correspondence, but the average ratio of the best and the average of second best correspondences across a set of images.  Rabin et al.\ \cite{rabin2009statistical} try to enhance descriptor matching by looking at the statistical distribution of local features in the matched images, and only return a match when such a correspondence would not occur by mere chance. Finally, Arnfred and Winkler generalize \emph{Ratio-Match} to make use of both of the matched images to provide a more accurate evaluation of the uniqueness of a match \cite{arnfred2013mirror}.

@@

In this chapter we will introduce the fundamentals of \emph{Fast-Match} and motivate the design choices as we go. The algorithm is made of 3 components; finding seeds, finding matches, and exploring for other places where matches might be. After discussing a few design choices in Section~\ref{considerations} we discuss each component individually. Seed matches are discussed in Section~\ref{seeds}, collecting matches in Section~\ref{collection} and exploring for new matches in Section~\ref{exploring}.

@@

If we set out to design a truly fast matching algorithm, we cannot rely just on optimizing the matching step once the descriptors from both images have been found and computed. As illustrated in Figure~\ref{timings} finding and computing descriptors can easily account for 80\% of the time spent matching for bigger images. For this reason \emph{Fast-Match} is designed to only compute features for the part of the image we hope to match.

@@

We introduce two variations of the algorithm, the a \emph{general} and the \emph{retrieval} approach. The \emph{retrieval} approach assumes that we know one of the images that we intend to match ahead of time and can do some off-line computations. For the \emph{general} approach we make those computations on the fly instead and do not assume that anything can be pre-computed. in In Figure~\ref{fig:match_example} we illustrate the result of matching \emph{Fast-Match} on two images compared with the result of \emph{Ratio-Match} for the same images.

\begin{figure}[tb]
    \centering
    \begin{subfigure}[t]{0.5\columnwidth}
        \includegraphics[width=1\columnwidth]{images/timing_large}
    \end{subfigure}%
    ~%\vspace{1.5 mm}
    \begin{subfigure}[t]{0.5\columnwidth}
        \includegraphics[width=1\columnwidth]{images/timing_medium}
    \end{subfigure}%
    \caption{The duration in seconds of different variations of \emph{Fast-Match} and \emph{Ratio-Match}. The top plot shows results for two 10~MPixel images two 3~Mpixel images is used for the comparison at the bottom. In both plots from left to right: 1.~The proposed \emph{Fast-Match} algorithm with one image pre-cached. 2.~The proposed \emph{Fast-Match} algorithm with no pre-cached images. 3-5.~ \emph{Ratio-Match} using respectively Flann KD-Tree, KGraph and Brute-Force to find nearest neighbours}
    \label{fig:timings}
\end{figure}

@@

\subsection{Initiating Seeds}
\label{seeds}
To obtain a set of initial seed matches several strategies can be used depending on the use case. For the case of matching a pair of two images we can get a rough group of seed matches by matching the thumbnails of the two images with for example \emph{Ratio-Match}. However if we were matching images in a series, such as frames from a movie we could instead make use of a subset of the matches from the last frame to seed \emph{Fast-Match} on the next.

@@

Lowe and others have shown that the distance between two SIFT descriptors is much less indicative of a true correspondence than the ratio between the best and second best match \cite{lowe2004sift,mikolajczyk2005performance,moreels2007evaluation,rabin2009statistical}. This ratio is more formally defined as follows: If we let $f_q$ be a feature in the \emph{query image} and $f_t, f_b$ be the two nearest neighbours of $f_q$ in the \emph{target image} then the ratio $r$ is defined as follows:
\begin{align*}
    r &= \text{r}(f_{q}, (f_t, f_b)) \\
        &= \frac{d(f_{q}, f_{t})}{d(f_{q}, f_{b})}.
\end{align*}
Here $d(f_i,f_j)$ is the distance between the features $f_i$ and $f_j$. For SIFT this is the euclidean distance. Using $r$ as measure for match confidence presumes that we expect a one-to-one matching of the features between the two images. Intuitively if we try to find a match for a feature $f_i$ in an image that does not have any true correspondences, then we would expect the two closest neighbours to be roughly equally well matched with $f_i$. For this reason we attribute high confidence to matches where the closest neighbour is dramatically closer to $f_i$ than the second closest neighbour and discard the rest.

We are faced with a problem when applying this technique to obtain the set of confidence scores $C_{qt}$ since for any match in $M_{qt}$ we only know the nearest neighbours amongst the features of $R_q$ and $R_t$. To get around this problem we either assume to know one of the images beforehand (the \emph{retrieval} approach) or in case we cannot make that assumption we compute the features of one of the images (the \emph{general} approach). For a given match between features $f_q$ and $f_t$ we can now find the second closest neighbour neighbour $f_b$ and calculate the confidence as $r = \frac{d(f_q, f_t)}{d(f_q,f_b)}$.

In the case that we know the image beforehand, we can further optimize this step. If we assume the \emph{target image} is known in advance we can approximate the ratio by letting $\hat r = \frac{d(f_t, f_q)}{d(f_t,f_b)}$. Here the feature $f_b$ is also part of the target image, which means we can pre-calculate $d(f_t, f_b)$ for all features in the cached \emph{target image} before we start matching.

@@

There are many possible heuristics for predicting possible matching regions including using local and global epipolar assumptions and partial graph isomorphisms. However for the sake of simplicity and speed we have chosen a straightforward approach based on weak angular assumptions and illustrated in Figure~\ref{fig:exploration}. Assuming that the \emph{target image} has been cached either in advance or before the matching step, the blue shaded areas show $R_q$ and $R_t$ for a given seed match. In the \emph{target image} we collect all features in a given radius while we compute features in the rectangular $R_q$ in the \emph{query image}. For performance reasons we compute all features in the blue square but match only the features inside the shaded area. Next a match is found in the collection step between $f_q$ and $f_t$. Based on the position of $f_q$ in $R_q$ we select three areas with potential for more matches. The center of each is matched with the center position $f_t$ to produce three seed matches for the next iteration.

In practice it is necessary that these squares overlap in order to detect features lying close to the edges. This incurs a bit of overhead which is minimized by only collecting features for groups of 9 squares. In order to avoid doing double work, quite a bit of care has to be expended making sure that results are properly cached. In Figure~\ref{fig:match_example} the areas across all iterations are indicated to illustrate a typical result.

\subsection{Computational Complexity}
\label{complexity}
The difference between the \emph{general} and \emph{retrieval} approach of \emph{Fast-Match} consists in whether we compute the features in the \emph{target image} (general) or if we presume the features are computed offline (retrieval). Computing the features is linear in the amount of features, while finding $d(f_q, f_b)$ for all $n$ features in the \emph{target image} can be done in $O(nlogn)$ using metric trees.

@@

We can provide an upper bound on the probability that a particular pair of images will exhibit a super-linear behaviour by observing that:
\begin{align}
    \mathbb{P}\left(\left| \frac{S_n}{n} - \mathbb{E}(X_t) \right| \le f(n)\right) &\le \frac{Var\left(\frac{S_n}{n}\right)}{(f(n))^2} & (chebyshev)\\
&\le \frac{6}{n(f(n))^2}
\end{align}

For larger values of $n$ it be becomes increasingly likely that \emph{Fast-Match} at worst will show linear performance in terms of the amount of target features.

The noteworthy part of this result is that for the \emph{retrieval} approach we can match two images in $O(n)$. Since there are no large constants involved this makes it possible to rapidly match very large images. In particular \emph{Fast-Match} is suited for cases where we are looking to match several large \emph{query images} to one \emph{target image}. In this case we can compute the features and distances of the \emph{target image} and then match each \emph{query image} in linear time.

\section{Experimental Setup}
\label{experiments}
% TODO: get exact number of image pairs
We evalute \emph{Fast-Match} over 3024 image pairs featuring various 3D objects at different angles. We compare the algorithm to the standard \emph{Ratio-Match}\cite{lowe2004sift} as well as the newer \emph{Mirror-Match}\cite{arnfred2013mirror}. These two algorithms were selected because they are magnitudes faster than the fastest geometric algorithms while at the same time providing robust matches.

@@

The 3D objects dataset by Moreels and Pietro \cite{moreels2007evaluation} allows us to experimentally compare matching algorithms over a large range of object and surface types rotated on a turnstile and photographed from every 5 degree turn.  15 objects from the dataset are shown in Figure~\ref{fig:3d_objects}.  We use images of 84 different objects under three different lighting conditions at 12 different angle intervals, conducting experiments with a total of 3024 image pairs.

@@

To validate matches, Moreels and Pietro propose a method using epipolar constraints \cite[p.266]{moreels2007evaluation}, which is outlined in Figure~\ref{fig:frog}.  According to their experiments, these constraints are able to identify true correspondences with an error rate of $2\%$. We use their proposed method to generate the ground truth for the evaluation of our framework.

\begin{figure}[htb]
    \centering
    \begin{subfigure}[t]{0.22\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_A}
        \caption{Query Image}
        \label{fig:frog_a}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad a 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.22\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_B}
        \caption{Auxiliary Image}
        \label{fig:frog_b}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad 
    %a blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.22\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C}
        \caption{Target Image}
        \label{fig:frog_c}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.22\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C_lines}
        \caption{Target Image with epipolar constraints}
        \label{fig:frog_c_lines}
    \end{subfigure}%
    \caption{Creating epipolar constraints based on three source images \cite{moreels2007evaluation}.  (a) \emph{Query image} marked with the position of the feature we are attempting to match. (b) Auxiliary image, taken at the same rotation as the \emph{query image} but from a higher elevation angle. The line going through the image is the epipolar line of the feature point in the \emph{query image}.  The markers indicate all feature points in the image found near the epipolar line.  In c) we show the \emph{target image} which in this case is rotated 45 degrees from the \emph{query image}.  The line overlaid on the image represents the epipolar line corresponding to the feature point shown in the \emph{query image}.  The markers indicate all feature points in the image found near the epipolar line. In d) we show the \emph{target image} again, this time overlaid with the epipolar lines corresponding to all the features shown in the \emph{auxiliary image}. A true correspondence should be found within a small distance of one of the intersections of the line in c) and the lines in d). In this particular case both feature points shown in c) and d) could possibly be a correct correspondence.
}%
    \label{fig:frog}%
\end{figure}%

To compute the total number of possible correspondences, we take each feature in a \emph{query image} and count how many of them have a feature in the \emph{target image} which would satisfy the epipolar constraints outlined above. When using this dataset, features with no correspondences were not included in the set of features for testing, so as to avoid matching non-moving background and foreground objects.

@@

\subsection{Configuration of \emph{Fast-Match}}
The central parameters of \emph{Fast-Match} are the confidence threshold, $\tau$ for selecting seed matches and the final confidence threshold applied to the total set of matches. For the experiments we have let $\tau = 0.9$ and created precision/recall plots by varying the confidence threshold over the final set of matches.

To achieve a good balance between speed and robustness we let the region in which we extract features be a square with a side length of 90 pixels. On empirical tests we found that anything smaller would result in decreased performance while much bigger Regions would decrease speed. The 90 pixel window is split in to nine smaller squares. When a seed match falls in any of these squares we match only the features within the smaller square. We let both the regions and the smaller squares overlap each other at all edges with 25 pixels in order to capture feature points lying close to an edge. For the image we have already cached we find all features within a radius of 50 pixels of the seed match.

While these parameters have worked well in practice, we are convinced that there exists other and better ways to implement \emph{Fast-Match} in particular for specific applications. We encourage experimentation and will happily assist with advice.

@@

@@

@@

% Include graph with different images as baseline set and the consequent matching quality

% - It's an extension of the journal submission in a sense. Instead of
%   using either the target or query image as baseline set we go a step
%   further and ask if we can use a general baseline image

% - Talk about metric trees and approximate nearest neighbours

% - Show figure that breaks down the time spent in different parts of a
% matching algorithm.

% - Introduce the two modes of matching. One using general background
%   info and one using the query image

In this paper we would like to introduce \emph{Fast-Match}, a matching algorithm designed to speed up

Matching image points is a crucial ingredient in almost all computer vision applications that deal with sparse local image features, such as image categorization \cite{bosch2008scene}, image stitching \cite{brown2007automatic}, object detection \cite{zhang2007local}, and near duplicate detection \cite{zhao2009scale}, to mention just a few examples.  All of these rely on accurately finding the correspondence(s) of a point on an object in a \emph{query image} given one or more \emph{target images} that might contain the same object.  In many applications the \emph{target images} have undergone transformations with respect to the \emph{query image}; in stereo vision, the viewpoint is different, while in object recognition and near duplicate  detection both the lighting and even the object itself may also be transformed.

In the literature two approaches to feature point matching have been pursued and later merged. The geometric approach tries to find unique keypoints in an image pair and matches them based on their location.  The descriptor centric approach on the other hand uses the local image information around the keypoint to create a descriptor and matches descriptors based on their similarity.

In the purely geometric approach, feature points are matched based on their position in the images. Scott and Longuet-Higgins \cite{scott1991algorithm} and Shapiro and Brady \cite{shapiro1992feature} introduced the use of spectral methods by deriving a coherent set of matches from the eigenvalues of the correspondence matrix. Other examples of this approach include \cite{sclaroff1995modal,carcassoni2003spectral}.

The descriptor centric approach on the other hand finds matches by pairing similar keypoints. The first examples of this approach used the correlation of the raw image data immediately surrounding the feature point \cite{deriche1994robust,baumberg2000reliable} to calculate this similarity. Later algorithms were enhanced by invariant feature descriptors as first introduced in \cite{schmid1997local} and later popularized by the work of Lowe introducing SIFT \cite{lowe2004sift} and Bay et al.\ introducing SURF \cite{bay2006surf}.

Many descriptor variations have since been proposed, and a few surveys have emerged comparing them. Mikolajczyk and Schmid \cite{mikolajczyk2005performance} evaluate 11 such feature descriptors using images of planar scenes.  Moreels and Perona \cite{moreels2007evaluation} evaluate 5 descriptors with different keypoint detectors using a large image set consisting of 3D models. Heinly et al.\ \cite{heinly2012comparative} take a closer look at binary features such as BRIEF \cite{calonder2010brief}, BRISK \cite{leutenegger2011brisk}, and ORB \cite{rublee2011orb}, comparing them to SIFT and SURF.

A straightforward way to find a set of correspondences using only feature points is to apply a threshold to the similarity measure of the feature vectors, accepting only correspondences that score above a certain level of similarity \cite{szeliski2010}. When we match images with the assumption that the correspondence between two feature points will be unique, we can further increase precision by only matching a feature point to its nearest neighbour in terms of descriptor similarity. Instead of thresholding based on similarity, Deriche et al.~\cite{deriche1994robust} and Baumberg~\cite{baumberg2000reliable} propose using the ratio of the similarity of the best to second best correspondence of a given point to evaluate how unique it is. Their finding has later been tested by several independent teams, all concluding that thresholding based on this ratio is generally superior to thresholding based on similarity \cite{lowe2004sift,mikolajczyk2005performance,moreels2007evaluation,rabin2009statistical}.

>>>>

%===========================================================
\title{Fast-Match: Fast and robust matching on large images}
\titlerunning{ACCV-14 submission ID \ACCV14SubNumber}
\authorrunning{ACCV-14 submission ID \ACCV14SubNumber}

@@

\begin{abstract}
    Both consumer cameras and camera phones produce images that often exceed 10 mega pixels. Yet computing and matching local image features in images of this size can easily take more than twenty seconds using fast matching algorithms. This is much too slow for interactive applications and much too expensive for large scale image operations. We introduce \emph{Fast-Match}, an algorithm designed to swiftly match large images without compromising on matching precision or recall. \emph{Fast-Match} derives its speed from only computing features in the parts of the image that can be confidently matched. We show that \emph{Fast-Match} is an order of magnitude faster than \emph{Ratio-Match} and demonstrate that \emph{Fast-Match} often doubles the precision on difficult to match image pairs at equal recall. In addition we prove that when one image is known in advanced, \emph{Fast-Match} can achieve linear performance matching in the amount of feature points.
\end{abstract}

\section{Introduction}
%
Whenever we match local image features we are faced with a choice between performance and precision. On one hand SIFT features proposed by Lowe \cite{lowe2004sift} have shown again and again to compare favorably to other local image descriptors especially under unrestrained conditions such as perspective change with non-planar scenes \cite{mikolajczyk2005performance,moreels2007evaluation,heinly2012comparative}. On the other, SIFT keypoints and descriptors are slow to compute, the main raison d'\^{e}tre for other local image features. In many applications of computer vision we would like the increase the computational performance in order to work on larger images, bigger image sets, at faster frame rates or with more limited hardware but we cannot give up the additional precision that SIFT affords us over other local features without negative effects for our application.

In this paper we introduce a matching algorithm designed to match features between two images only in image areas that are likely to correspond. This approach is much faster than traditional methods because there is no need to compute descriptors for areas in the image that are not matched. We provide two variations of the algorithm: a \emph{general} and a \emph{retrieval} variation. The \emph{general} variation functions as a traditional matching algorithm and matches two unknown images albeit faster and more robustly than conventional methods. The \emph{retrieval} variation on the other hand assumes that we know one of the images we intend to match beforehand but given this trade-off this variation offers performance a magnitude faster than existing matching methods without compromising on accuracy. We will unimaginatively refer to the proposed algorithm as \emph{Fast-Match} in this paper and make clear from the context whether this refers to the \emph{retrieval} or \emph{general} variation. As an aside we adopt an artificial distinction between the words \emph{match} and \emph{correspondence} to make it clear when a match between two images is correct. The word \emph{match} will be used to denote a result from a matching algorithm that may or may not be correct while a \emph{correspondence} refers to an actual visual correspondence between two points in two images.

@@

Efforts to reduce the computational costs of finding nearest neighbours to feature points has largely been focused on metric trees. Na\"ively the set of nearest neighbours between features in two images can computed by brute force in $O(n^2)$ where $n$ is the number of features in the two images. Typically a three mega pixel image contains anywhere from 500 to 5000 feature points. When SIFT was originally published Lowe proposed using the Best-Bin-First method to approximately search for nearest neighbours \cite{beis1997shape,lowe1999object}. This reduces the computational complexity to $O(n\log n)$ but even approximate metric trees are hard pressed to compete with brute force due to the high dimensionality of SIFT and the constant costs incurred with constructing and searching in a metric tree. Later work by Muja and Lowe has focused on improving approximate nearest neighbour searches by using several KD-Trees simultaneously while optimizing the tree structure using K-Means to cluster similar features \cite{muja2009fast}. Recent work on knn-graphs shows a lot of promise for high dimensional cases \cite{dong2011efficient}. We later review these improvements and their effect on efficiently matching large scale images.

@@

Brown and Lowe \cite{brown2005multi} extend ratio match to deal with a set of images by using not the ratio of the best and second best correspondence, but the average ratio of the best and the average of second best correspondences across a set of images.  Rabin et al.\ \cite{rabin2009statistical} try to enhance descriptor matching by looking at the statistical distribution of local features in the matched images, and only return a match when such a correspondence would not occur by mere chance. Finally, Arnfred et al.\ generalize \emph{Ratio-Match} to make use of both of the matched images to provide a more accurate evaluation of the uniqueness of a match \cite{arnfred2013mirror}.

@@

In this chapter we will introduce the fundamentals of \emph{Fast-Match} and motivate the design choices as we go. The algorithm is made of 3 components; finding seeds, finding matches, and exploring for other places where matches might be.

@@

If we set out to design a truly fast matching algorithm, we cannot rely just on optimizing the matching step once the descriptors from both images have been found and computed. As illustrated in Figure~\ref{fig:timings} finding and computing descriptors can easily account for 80\% of the time spent matching for bigger images. For this reason \emph{Fast-Match} is designed to only compute features for the part of the image we hope to match.

@@

We introduce two variations of the algorithm, the a \emph{general} and the \emph{retrieval} variation. The \emph{retrieval} variation assumes that we know one of the images that we intend to match ahead of time and can do some off-line computations. For the \emph{general} variation we make those computations on the fly instead and do not assume that anything can be pre-computed. in In Figure~\ref{fig:match_example} we illustrate the result of matching \emph{Fast-Match} on two images compared with the result of \emph{Ratio-Match} for the same images.

\begin{figure}[tb]
    \centering
    \includegraphics[width=1\columnwidth]{images/timings}
    \caption{The duration in seconds of different variations of \emph{Fast-Match} and \emph{Ratio-Match}. The top plot shows results for two 10~MPixel images two 3~Mpixel images is used for the comparison at the bottom. In both plots from left to right: 1.~The proposed \emph{Fast-Match} algorithm with one image pre-cached. 2.~The proposed \emph{Fast-Match} algorithm with no pre-cached images. 3-5.~ \emph{Ratio-Match} using respectively Flann KD-Tree, KGraph and Brute-Force to find nearest neighbours}
    \label{fig:timings}
\end{figure}

@@

\subsection{Initiating Seeds}
\label{seeds}
To obtain a set of initial seed matches several strategies can be used depending on the use case. For the case of matching a pair of two images we can get a rough group of seed matches by matching the thumbnails of the two images with for example \emph{Ratio-Match}. However if we were matching images in a series, such as frames from a movie we could instead make use of a subset of the matches from the last frame to seed \emph{Fast-Match} on the next. In practice we have found it efficient to resize both images to thumbnails of 300 by 300 pixels and use ratio match to obtain a set of matches and ratios that we then threshold with the confidence threshold $\tau$ to obtain the initial seed matches.

@@

Lowe and others have shown that the distance between two SIFT descriptors is much less indicative of a true correspondence than the ratio between the best and second best match \cite{lowe2004sift,mikolajczyk2005performance,moreels2007evaluation,rabin2009statistical}. This ratio is more formally defined as follows: If we let $f_q$ be a feature in the \emph{query image} and $f_t, f_b$ be the two nearest neighbours of $f_q$ in the \emph{target image} then the ratio $r$ is defined as follows:
\begin{align*}
    r &= \text{r}(f_{q}, (f_t, f_b)) \\
        &= \frac{d(f_{q}, f_{t})}{d(f_{q}, f_{b})}.
\end{align*}
Here $d(f_i,f_j)$ is the distance between the features $f_i$ and $f_j$. For SIFT this is the euclidean distance. The distance between two feature points is greater when the feature points resemble each other less. For this reason we have higher confidence in a match with a low $r$-value. Using $r$ as measure for match confidence presumes that we expect each feature in the \emph{target image} to have at most one true correspondence in the \emph{target image}. Intuitively if we try to find a match for a feature $f_i$ in an image that does not have any true correspondences, then we would expect the two closest neighbours to be roughly equally well matched with $f_i$. For this reason we attribute high confidence to matches where the closest neighbour is dramatically closer to $f_i$ than the second closest neighbour and discard the rest.

We are faced with a problem when applying this technique to obtain the set of confidence scores $C_{qt}$ since for any match in $M_{qt}$ we only know the nearest neighbours amongst the features of $R_q$ and $R_t$. To get around this problem we either assume to know one of the images beforehand (the \emph{retrieval} variation) or in case we cannot make that assumption we compute the features of one of the images (the \emph{general} variation). For a given match between features $f_q$ and $f_t$ we can now find the second closest neighbour neighbour $f_b$ and calculate the confidence as $r = \frac{d(f_q, f_t)}{d(f_q,f_b)}$.

In the case that we know the image beforehand, we can further optimize this step. If we assume the \emph{target image} is known in advance we can approximate the ratio by letting $\hat r = \frac{d(f_t, f_q)}{d(f_t,f_b)}$. Here the feature $f_b$ is also part of the target image, which means we can pre-calculate $d(f_t, f_b)$ for all features in the cached \emph{target image} before we start matching. When these distances are tied to their corresponding feature in the target image, it becomes trivial to calculate $\hat r$.

@@

There are many possible heuristics for predicting possible regions including using local and global epipolar assumptions and partial graph isomorphisms. However for the sake of simplicity and speed we have chosen a straightforward approach based on weak angular assumptions and illustrated in Figure~\ref{fig:exploration}. Assuming that the \emph{target image} has been cached either in advance or before the matching step, the blue shaded areas show $R_q$ and $R_t$ for a given seed match. In the \emph{target image} we collect all features in a given radius while we compute features in the rectangular $R_q$ in the \emph{query image}. For performance reasons we compute all features in the blue square but match only the features inside the shaded area. Next a match is found in the collection step between $f_q$ and $f_t$. Based on the position of $f_q$ in $R_q$ we select three areas with potential for more matches. The center of each is matched with the center position $f_t$ to produce three seed matches for the next iteration.

In practice it is necessary that these squares overlap in order to detect features lying close to the edges. This incurs a bit of overhead which is minimized by only collecting features for groups of 9 squares. In order to avoid computing the same matches or features twice, quite a bit of care has to be expended making sure that results are properly cached. A matrix containing `bins' of features can conveniently be used to store features from different image regions. Similarly a hash-set is suitable for keeping track of which regions have already been matched and which matches have already been found.

\subsection{Computational Complexity}
\label{complexity}
The difference between the \emph{general} and \emph{retrieval} variation of \emph{Fast-Match} consists in whether we compute the features in the \emph{target image} (general) or if we presume the features are computed offline (retrieval). Computing the features is linear in the amount of features, while finding $d(f_q, f_b)$ for all $n$ features in the \emph{target image} can be done in $O(n\log n)$ using metric trees.

@@

For the above case it is theoretically possible that we for every one of the $n$ features end up considering $kn$ seed matches for some constant $k$, i.e. a quadratic behaviour. We can provide an upper bound on the probability of this event:
\begin{align}
    \label{bound}
    \mathbb{P}\left(\left| \frac{S_n}{n} - \mathbb{E}(X_t) \right| \ge kn\right) &\le \frac{Var\left(\frac{S_n}{n}\right)}{k^2n^2} & (chebyshev)\\
&\le \frac{6}{k^2n^3}
\end{align}

It is clear that for any constant $k$ we can pick a number of features $n$ which renders the likelihood of a quadratic behaviour infinitesimal. In essence the larger $n$ becomes, the less likely \emph{Fast-Match} is to exhibit super linear behaviour.

The noteworthy part of this result is that for the \emph{retrieval} variation we can match two images in $O(n)$. Since there are no large constants involved this makes it possible to rapidly match very large images. In particular \emph{Fast-Match} is suited for cases where we are looking to match several large \emph{query images} to one \emph{target image}. In this case we can compute the features and distances of the \emph{target image} and then match each \emph{query image} in linear time. In contrast the general variation still has a complexity of $O(n \log n)$ due to the necessity of calculating distances online.

\section{Experimental Setup}
\label{experiments}
% TODO: get exact number of image pairs
We evalute \emph{Fast-Match} over 3024 image pairs featuring various 3D objects at different angles to measure precision and recall, as well as two pairs of images with high pixel counts to measure performance in terms of speed. We compare the algorithm to the standard \emph{Ratio-Match}\cite{lowe2004sift} as well as the newer \emph{Mirror-Match}\cite{arnfred2013mirror}. These two algorithms were selected because they are magnitudes faster than the fastest geometric algorithms while at the same time providing robust matches.

@@

% TODO: recompute number of image pairs
The 3D objects dataset by Moreels and Pietro \cite{moreels2007evaluation} allows us to experimentally compare matching algorithms over a large range of object and surface types rotated on a turnstile and photographed from every 5 degree turn.  15 objects from the dataset are shown in Figure~\ref{fig:3d_objects}.  We use images of 84 different objects under three different lighting conditions at 12 different angle intervals, conducting experiments with a total of 3024 image pairs.

@@

To validate matches, Moreels and Pietro propose a method using epipolar constraints \cite[p.266]{moreels2007evaluation}.  According to their experiments, these constraints are able to identify true correspondences with an error rate of $2\%$. We use their proposed method to generate the ground truth for the evaluation of our framework.

To compute the total number of possible correspondences, we take each feature in a \emph{query image} and count how many of them have a feature in the \emph{target image} which would satisfy the epipolar constraints. When using this dataset, features with no correspondences were not included in the set of features for testing, so as to avoid matching non-moving background and foreground objects.

@@

\subsection{Configuration of \emph{Fast-Match}}
The central parameters of \emph{Fast-Match} are the confidence threshold, $\tau$ for selecting seed matches and the final confidence threshold applied to the total set of matches. For the experiments on the 3D Objects we let $\tau = 0.9$ and created precision/recall plots by varying the confidence threshold over the final set of matches.

To achieve a good balance between speed and robustness we let the region in which we extract features be a square with a side length of 90 pixels. On empirical tests we found that anything smaller would result in decreased performance while much bigger regions would decrease speed. The 90 pixel window is split in to nine smaller squares. When a seed match falls in any of these squares we match only the features within the smaller square. We let both the regions and the smaller squares overlap each other at all edges with 25 pixels in order to capture feature points lying close to an edge. For the image we have already cached we find all features within a radius of 50 pixels of the seed match.

While these parameters have worked well in practice, we are sure that there exists other and better ways to implement \emph{Fast-Match} in particular for specific applications. We encourage experimentation and will happily assist with advice.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1\columnwidth]{images/grid_match_all}
\caption{Results for the 3d objects dataset. Each plot contains data accumulated from 84 objects photographed under 3 different lighting conditions}
\label{fig:all_objects}
\end{figure}

At low angular differences all algorithms show similar performance at low recall, however \emph{Fast-Match} is superior to \emph{Ratio-Match} and \emph{Mirror-Match} at higher recall, more than doubling the precision at similar recall levels. This picture remains the same at higher angular differences, but with a higher performance gab between \emph{Ratio-Match} and the other algorithms at low recall while more than doubling precision at higher recalls. Finally for image pairs with angular differences over 50$^0$ the performance of all algorithms decline with \emph{Fast-Match} still showing better results for the most part.

The high recall rates of \emph{Fast-Match} are partially explained by the fact that we can allow our confidence thresholds to be more lenient when we already know that we are likely to find correct correspondences in the regions that we are matching. However lenient thresholds can easily impact the matching speed, so in practical situations we will usually have to choose between matching fast and precise or slower with high recall.

Figure~\ref{fig:match_example} and \ref{fig:timings} demonstrate the speed of \emph{Fast-Match} on image sizes as is typically produced by modern consumer cameras and camera phones. In Figure~\ref{fig:timings} we compare the speed of the two variations of \emph{Fast-Match} to different variations of \emph{Ratio-Match} on two image pairs, one with two images of 10 mega pixel and another where the same image pair has been scaled to 3 mega pixels. For the larger images we see the clearest difference in performance. The retrieval variant of \emph{Fast-Match} matches the image pairs in around one second while a retrieval variant of \emph{Ratio-Match} with precomputed features spends eight seconds on the same task. This result is roughly equal to the time the general variation of \emph{Fast-Match} spends matching the two images without any pre-computations. For the nearest neighbor search we note that for large images there are very substantial speed gains to be had by choosing the right algorithm for finding nearest neighbors, with Muja et al.'s Flann matcher being vastly superior to brute force and KGraph.

For the 3 mega pixel picture the pattern repeats itself, although with smaller margins separating \emph{Fast-Match} and \emph{Ratio-Match}. When images are smaller, computations like obtaining the initial seed matches are comparatively more costly.

Figure~\ref{fig:match_example} shows the result from matching the pair of three mega pixel images using \emph{Fast-Match} and \emph{Ratio-Match}. An almost identical result can be obtained using the pair of 10 mega pixel pictures instead. In the figure we see how \emph{Fast-Match} only ends up matching the part of the images that are likely to have correspondences. This highlights both a strength and a weakness of \emph{Fast-Match}. It is this myopic matching approach that yields \emph{Fast-Match} its speed and precision, but \emph{Fast-Match} will not be the best candidate for almost identical images where \emph{Fast-Match}'s step by step approach incurs too much overhead to compete in speed with \emph{Ratio-Match}.

@@

We have introduced \emph{Fast-Match} in two forms, a general and a retrieval variation. The retrieval variation assumes that we know one of the images that we are matching ahead of time in order to pre-cache feature points. The general variation has no assumptions and performs the pre-caching of feature points online instead. 
We have proven that while the general variation has a computational complexity of $O(n \log n)$ with $n$ being the number of feature points, the retrieval variation will on average run in $O(n)$.
From experiments on large images we have shown that \emph{Fast-Match} can be a magnitude faster than \emph{Ratio-Match} while providing comparable results. We went on to evaluate \emph{Fast-Match} to \emph{Ratio-Match} and \emph{Mirror-Match} using images of rotated 3D objects and demonstrated that \emph{Fast-Match} outperforms the other algorithms significantly over 3024 image pairs and in most cases double the matching precision at similar recall rates.
\emph{Fast-Match} is particularly applicable in cases where we are interested in matching one image to multiple large images, in which case we can quickly pre-cache the image and use this data for each of the other images.

@@
